How fluid-mechanical erosion creates anisotropic porous media

### 2022-12-01 docs/PhysicsPaper/StokesErosion.tex
NM

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% COMMANDS
% Basic editing
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\tocite}{ {\color{blue}(to cite)} }
\newcommand{\nick}[1]{ {\color{red} #1} }
% For real and imaginary, could use \Re or \Im, or \mathcal{R}, or \text{Re}
\newcommand{\Real}{\Re}
\newcommand{\Imag}{\Im}

@@

\begin{abstract}
Using a Cauchy integral formulation of the boundary integral equations, we simulate the erosion of a porous medium comprised of up to 100 solid bodies embedded in a Stokes flow. The grains of the medium are resolved individually and erode under the action of surface shear stress. Through nonlinear feedback with the surrounding flow fields, microscopic changes in grain morphology give way to larger-scale features in the medium such as channelization. The Cauchy-integral formulation and associated quadrature formulas enable us to resolve dense configurations of nearly contacting bodies.
We observe substantial anisotropy to develop over the course of erosion; that is, the configurations that result from erosion generally permit flow in the longitudinal direction more easily than in the transverse direction by up to a factor of six. These results suggest that the erosion of solid material from groundwater flows may contribute to previously observed anisotropy of natural porous media.
\end{abstract}
\maketitle

@@

Our method merges highly-efficient and highly-accurate boundary-integral equation (BIE) methods~\cite{baker1986boundary, moore2007evaluation, gray2019boundary} with stable interface evolution methods~\cite{hou-low-she1994, Moore2013} to simulate the erosion of dense suspensions of solid bodies in the Stokes flow regime relevant for groundwater-flow applications~\cite{quaife2018boundary, chiu2020viscous}. Originally inspired by related work in the high-Reynolds-number regime~\cite{Ristroph2012, Moore2013, Huang2015, MooreCPAM2017}, our method is documented, validated, benchmarked in~\cite{quaife2018boundary}, and can simulate the erosion of $O(100)$ solid bodies. The more recent Cauchy reformulation of the BIE and the associated quadrature formulas allow us to resolve points of near contact between bodies, thus enabling high-fidelity simulation of dense suspensions of erodable bodies~\cite{chiu2020viscous}. Though the governing fluid-flow equations (Stokes) are linear, the nonlinear feedback between evolving microstructure and flow gives rise to highly complex and anisotropic configurations.

@@

The embedded bodies may erode in response to the shear stresses induced by the intervening fluid flow. Erosion typically occurs over much longer timescales than the fluid flow, permitting a {\em quasi-steady} approximation. In this approximation, the configuration of bodies is held fixed in order to compute the {\em steady} Stokes flow determined by~\eqref{eqn:StokesEq}, and then this flow field determines the stresses acting to erode each body. We employ an erosion law in which the local rate of material loss is linearly proportional to the magnitude of the shear stress $\tau$ acting on the surface~\cite{Ristroph2012, Moore2013, Mitchell2016, MooreCPAM2017, HewettJFS2017, quaife2018boundary, chiu2020viscous}. The material loss gives rise to an inward velocity of the solid surface, $\Vn$, pointing in the direction normal to the surface. The erosion law is thus expressed as
\begin{align}
\Vn = \CE \, \abs{\tau}, 
	&\hspace{20pt} \xx \in \gamma, &&\mbox{\em erosion model}, 
\label{Vn0} \\
\tau = -\mu \left( ( \nabla \uu + \nabla \uu^T) \nn \right) \cdot \ss,
	&\hspace{20pt} \xx \in \gamma, &&\mbox{\em shear stress}.
\label{tau0}
\end{align}
where $\nn$ is the unit normal vector pointing into each body, $\ss$ is the unit tangent vector pointing in the counterclockwise direction, and $\CE$ is a material-dependent erosion constant.

% SNIPPETS 
%The strength of $\UU$ is adjusted at each time step to achieve a constant pressure drop across the channel, motivated by the geological situation of a porous medium connecting two regions of  fixed hydraulic heads.

%to transition to the more complex flow intervening between the bodies.

@@

%-----------------------------------------------------------------------------------------------%
\subsection{Interface evolution}
%-----------------------------------------------------------------------------------------------%

@@

The late emergence of substantial configurational anisotropy may seem
like a spurious effect that is particular to the simulation shown in
Figure~\ref{fig3}. However, statistical analysis of many simulations will show this feature to emerge robustly in simulations featuring a relatively large number of bodies $M \ge 60$.

@@

\subsection{Statistics of ensembles of simulations}

>>>>

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% COMMANDS
% Basic editing
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\tocite}{ {\color{blue}(to cite)} }
\newcommand{\nick}[1]{ {\color{red} #1} }
\newcommand{\edit}[1]{ {\color{red} #1} }

% For real and imaginary, could use \Re or \Im, or \mathcal{R}, or \text{Re}
\newcommand{\Real}{\Re}
\newcommand{\Imag}{\Im}

@@

\begin{abstract}
Using a Cauchy integral formulation of the boundary integral equations, we simulate the erosion of a porous medium comprised of up to 100 solid bodies embedded in a \edit{two-dimensional} Stokes flow. The grains of the medium are resolved individually and erode under the action of surface shear stress. Through nonlinear feedback with the surrounding flow fields, microscopic changes in grain morphology give way to larger-scale features in the medium such as channelization. The Cauchy-integral formulation and associated quadrature formulas enable us to resolve dense configurations of nearly contacting bodies.
We observe substantial anisotropy to develop over the course of erosion; that is, the configurations that result from erosion generally permit flow in the longitudinal direction more easily than in the transverse direction by up to a factor of six. These results suggest that the erosion of solid material from groundwater flows may contribute to previously observed anisotropy of natural porous media.
\end{abstract}
\maketitle

@@

Our method merges highly-efficient and highly-accurate boundary-integral equation (BIE) methods~\cite{baker1986boundary, moore2007evaluation, gray2019boundary} with stable interface evolution methods~\cite{hou-low-she1994, Moore2013} to simulate the erosion of dense suspensions of solid bodies in \edit{two-dimensional Stokes flows. We employ the Stokes equations because it is the regime relevant for groundwater applications~\cite{quaife2018boundary, chiu2020viscous}}. Originally inspired by related work in the high-Reynolds-number regime~\cite{Ristroph2012, Moore2013, Huang2015, MooreCPAM2017}, our method is documented, validated, benchmarked in~\cite{quaife2018boundary}, and can simulate the erosion of $O(100)$ solid bodies. The more recent Cauchy reformulation of the BIE and the associated quadrature formulas allow us to resolve points of near contact between bodies, thus enabling high-fidelity simulation of dense suspensions of erodable bodies~\cite{chiu2020viscous}. Though the governing Stokes equations are linear, the nonlinear feedback between evolving microstructure and flow gives rise to highly complex and anisotropic configurations.

@@

The embedded bodies may erode in response to the shear stresses induced by the intervening fluid flow. Erosion typically occurs over much longer timescales than the fluid flow, permitting a {\em quasi-steady} approximation. In this approximation, the configuration of bodies is held fixed in order to compute the {\em steady} Stokes flow determined by~\eqref{eqn:StokesEq}, and then this flow field determines the stresses acting to erode each body. We employ an erosion law in which the local rate of material loss is linearly proportional to the magnitude of the shear stress $\tau$ acting on the surface~\cite{Ristroph2012, Moore2013, Mitchell2016, MooreCPAM2017, HewettJFS2017, quaife2018boundary, chiu2020viscous}. The material loss gives rise to an inward velocity of the solid surface, $\Vn$, pointing in the direction normal to the surface. The erosion law is thus expressed as
\begin{align}
\Vn &= \CE \, \abs{\tau}, 
	\hspace{88pt} \xx \in \gamma, &&\mbox{\em erosion model}, 
\label{Vn0} \\
\tau &= -\mu \left( ( \nabla \uu + \nabla \uu^T) \nn \right) \cdot \ss,
	\hspace{20pt} \xx \in \gamma, &&\mbox{\em shear stress}.
\label{tau0}
\end{align}
where $\nn$ is the unit normal vector pointing into each body, $\ss$ is the unit tangent vector pointing in the counterclockwise direction, and $\CE$ is a material-dependent erosion constant.
\edit{For numerical stability, we will modify this law in Section \ref{sec:int_evo} to include a smoothing term that depends on the local curvature of each body.}

@@

%-----------------------------------------------------------------------------------------------%
\subsection{Interface evolution}
\label{sec:int_evo}
%-----------------------------------------------------------------------------------------------%

@@

The late emergence of substantial configurational anisotropy may seem like a spurious effect that is particular to the simulation shown in Figure~\ref{fig3}. 
\edit{The statistical analysis presented in Section \ref{sec:stat_sims}, however, reveals this feature to emerge robustly across many simulations as long as the number of bodies is relatively large, $M \ge 60$.}

@@

\subsection{Statistics of ensembles of simulations}
\label{sec:stat_sims}
